---
title: "Introduction to psymisc"
author: "Daniel Gromer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to psymisc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, warning=FALSE, echo=FALSE}
library(psymisc)
library(dplyr, warn.conflicts = FALSE)
library(ez)
library(tidyr)
```

## Descriptive statistics

When beginning to analyze a data set, it's usually a good idea to take a look at descriptive statistics. Base R offers the `aggregate` function for computing summary statistics (e.g. mean, standard deviation) of factorial data.

```{r}
aggregate(trait_anx ~ group + gender, hquest, FUN = "mean")
```

Computing a second statistic needs to be done using a second call to `aggregate`:

```{r}
aggregate(trait_anx ~ group + gender, hquest, FUN = "sd")
```

So, a shortcoming of this function is that it does not allow you to compute multiple statistics at once.
The [dplyr](https://cran.r-project.org/web/packages/dplyr/) package introduced another to compute summary statistics of grouped data with the functions `group_by` and `summarise`. (If you don't work with dplyr yet, you should definitely check it out: [dplyr vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/introduction.html))

```{r}
hquest %>%
  group_by(group, gender) %>%
  summarise(mean = mean(trait_anx), sd = sd(trait_anx))
```

The function `ds` in psymisc is a wrapper for these two functions and provides an easy formula interface for descriptive statistics:

```{r}
ds(hquest, trait_anx ~ group + gender)
```

Note that `ds` calculates the mean and standard error of the mean by default. This behavior can be changed by using the `funs` argument:

```{r}
ds(hquest, trait_anx ~ group + gender, funs = c("n", "mean", "sd", "moe"))
```

The `fplot` function can be used to get a graphical overview of descriptive statistics:

```{r}
fplot(hquest, acrophobia ~ group + gender)
# With SD
fplot(hquest, acrophobia ~ group + gender, error = "sd")
```

